---
title: "Tonmya_Prescriptions"
format: pdf
editor: source
date: "`r Sys.Date()`"
editor_options: 
  chunk_output_type: console
---


```{r}
#| echo: false
#| warning: false
library(gt)
library(knitr)
library(ggplot2)
library(tidyr)
library(dplyr)
library(lubridate)
```

## Some simple models based on Tonmya sales

What are we doing here? 
- Just trying to get a grasp on the rate of increase and projecting out if that rate continues. It is not necessarily realistic to take it out as far as projections provide. Nothing here should be regarded as investment advice. I am not a financial analyst or advisor.

```{r data}
#| echo: false
sales <- c(2, 16,54,93,179,268,212,187,264,324,367, 352, 451, 598, 639, 642, 750, 755)
scripts <- c(2,12,35,60,126,154,127,108,163,193,219, 205, 285, 371, 384, 401, 462, 483)
new_scripts <-  c(2, 12, 35, 60, 126, 154, 122, 97, 140, 170, 201, 185, 249, 327, 341, 347, 365, 384)
start_date <- as.Date("2025-11-14")
weekly_seq <- seq(start_date, by = "week", length.out = 52)
```

## Sales Projection - at weekly growth rate - *exponential model*

```{r sales exponential model}
#| echo: false
#| warning: false

df_sales <- data.frame(sales, "wk" = 1:length(sales))
mod_sales_exp <- nls(sales ~ a * (1 + r)^wk,
           data = df_sales,
           start = list(a = 17, r = 0))
summary(mod_sales_exp)

df_exp <- data.frame("wk" = 1:52, 
           "sales_reported" = c(df_sales$sales, rep(NA, 52 - length(df_sales$sales))),
           "exponential_model" = round(summary(mod_sales_exp)$coefficients[1] *
              (1 + summary(mod_sales_exp)$coefficients[2])^(1:52)),
            "yearly_rate_predicted_M" = 52 / 1e3 * summary(mod_sales_exp)$coefficients[1] *
              (1 + summary(mod_sales_exp)$coefficients[2])^(1:52))

gt(df_exp[1:(length(sales) + 4), ])  |>
  fmt_currency(columns = 2, decimals = 0, currency = "USD") |>
  fmt_currency(columns = 3, decimals = 0, currency = "USD") |>
  fmt_currency(columns = 4, decimals = 1, currency = "USD") |>
  cols_label(sales_reported = "sales, 000") |>
  cols_label(exponential_model = "exp, 000") |>
  cols_label(yearly_rate_predicted_M = "annual, M") |>
  tab_header(
    title = md("Exponential Model"))


df_exp[1:(length(sales) + 4), 1:3] %>% 
  pivot_longer(2:3) %>% 
  ggplot(., aes(wk, value, colour = name)) + geom_point() + geom_line() +
  ylab("$, 000") +
    ggtitle("Tonmya sales with exponential model")
#clean up
rm(df_exp)

```
## Sales Projection - at weekly rate - *linear model*

```{r sales linear model}
#| echo: false
#| warning: false

mod_sales_lin <- lm(sales ~ wk, data = df_sales)

summary(mod_sales_lin)
plot(mod_sales_lin)

df_lin <- data.frame("wk" = 1:52, 
           "sales_reported" = c(df_sales$sales, rep(NA, 52 - length(df_sales$sales))),
           "linear_model" = summary(mod_sales_lin)$coefficients[2] * 1:52,
            "yearly_rate_predicted_M" = 52 / 1e3 * summary(mod_sales_lin)$coefficients[2] * 1:52)

gt(df_lin[1:(length(sales) + 4), ]) |>
  fmt_number(columns = c(3, 4), decimals = 1) |>
  cols_label(sales_reported = "sales, 000") |>
  cols_label(linear_model = "exp, 000") |>
  cols_label(yearly_rate_predicted_M = "annual, M") |>
  tab_header(title = md("Linear Model"))

df_lin[1:(length(sales) + 4), 1:3] %>% 
  pivot_longer(2:3) %>% 
  ggplot(., aes(wk, value, colour = name)) + geom_point() + geom_line() +
  ylab("$, 000") +
    ggtitle("Tonmya sales with linear model")

#clean up
rm(df_lin)

```

## Scripts Projection - at weekly growth rate - *exponential model*


```{r scripts exponential model}
#| echo: false
#| warning: false
df_scripts <- data.frame(scripts, "wk" = 1:length(scripts))
mod_scripts_exp <- nls(scripts ~ a * (1 + r)^wk,
           data = df_scripts,
           start = list(a = 17, r = 0))
summary(mod_scripts_exp)

df_exp <- data.frame("wk" = 1:52, 
           "scripts_reported" = c(df_scripts$scripts, rep(NA, 52 - length(df_scripts$scripts))),
           "exponential_model" = round(summary(mod_scripts_exp)$coefficients[1] *
              (1 + summary(mod_scripts_exp)$coefficients[2])^(1:52)),
            "yearly_rate_predicted_M" = 52 / 1e3 * summary(mod_scripts_exp)$coefficients[1] *
              (1 + summary(mod_scripts_exp)$coefficients[2])^(1:52))

gt(df_exp[1:(length(scripts) + 4), ]) |>
  cols_label(scripts_reported = "total reported, 000") |>
  cols_label(exponential_model = "exp predicted, 000") |>
  cols_label(yearly_rate_predicted_M = "annual, M") |>
  tab_header(
    title = md("Scripts, Exponential Model"))


df_exp[1:(length(scripts) + 4), 1:3] %>% 
  pivot_longer(2:3) %>% 
  ggplot(., aes(wk, value, colour = name)) + geom_point() + geom_line() +
  ylab("total scripts") +
    ggtitle("Tonmya scripts with exponential model")
```
## scripts Projection - at weekly rate - *linear model*

```{r scripts linear model}
#| echo: false
#| warning: false

mod_scripts_lin <- lm(scripts ~ wk, data = df_scripts)

summary(mod_scripts_lin)
plot(mod_scripts_lin)

df_lin <- data.frame("wk" = 1:52, 
           "scripts_reported" = c(df_scripts$scripts, rep(NA, 52 - length(df_scripts$scripts))),
           "linear_model" = summary(mod_scripts_lin)$coefficients[2] * 1:52,
            "yearly_rate_predicted_M" = 52 / 1e3 * summary(mod_scripts_lin)$coefficients[2] * 1:52)

gt(df_lin[1:(length(scripts) + 4), ]) |>
  cols_label(scripts_reported = "total reported, 000") |>
  cols_label(linear_model = "linear predicted, 000") |>
  cols_label(yearly_rate_predicted_M = "annual, M") |>
  tab_header(title = md("Linear Model"))

df_lin[1:(length(scripts) + 4), 1:3] %>% 
  pivot_longer(2:3) %>% 
  ggplot(., aes(wk, value, colour = name)) + geom_point() + geom_line() +
  ylab("total scripts") +
    ggtitle("Tonmya scripts with linear model")
```
