---
title: "Tonmya_Prescriptions"
format: pdf
editor: source
date: "`r Sys.Date()`"
editor_options: 
  chunk_output_type: console
---


```{r}
#| echo: false
#| warning: false
library(gt)
library(knitr)
library(ggplot2)
library(tidyr)
library(dplyr)
library(lubridate)
```

## Some simple models based on Tonmya sales

What are we doing here? 
- Just trying to get a grasp on the rate of increase and projecting out if that rate continues. It is not necessarily realistic to take it out as far as projections provide. Nothing here should be regarded as investment advice. I am not a financial analyst or advisor.

```{r data}
#| echo: false
sales <- c(2, 16,54,93,179,268,212,187,264,324,367, 352, 451, 598, 639, 642, 750, 755, 935, 799, 811, 829, 950, 994)
scripts <- c(2,12,35,60,126,154,127,108,163,193,219, 205, 285, 371, 384, 401, 462, 483, 572, 525, 515, 548, 580, 616)
new_scripts <-  c(2, 12, 35, 60, 126, 154, 122, 97, 140, 170, 201, 185, 249, 327, 341, 347, 365, 384, 465, 401, 397, 404, 428, 449)
start_date <- as.Date("2025-11-14")
weekly_seq <- seq(start_date, by = "week", length.out = 52)
```

## Sales Projection - at weekly growth rate - *exponential model*

```{r sales exponential model}
#| echo: false
#| warning: false

df_sales <- data.frame(sales, wk = 1:length(sales), "week" = weekly_seq[1:length(sales)])
mod_sales_exp <- nls(sales ~ a * (1 + r)^wk,
           data = df_sales,
           start = list(a = 17, r = 0))
summary(mod_sales_exp)

df_exp <- data.frame("wk" = 1:52, weekly_seq, 
           "sales_reported" = c(df_sales$sales, rep(NA, 52 - length(df_sales$sales))),
           "exponential_model" = round(summary(mod_sales_exp)$coefficients[1] *
              (1 + summary(mod_sales_exp)$coefficients[2])^(1:52)),
            "yearly_rate_predicted_M" = 52 / 1e3 * summary(mod_sales_exp)$coefficients[1] *
              (1 + summary(mod_sales_exp)$coefficients[2])^(1:52))

gt(df_exp[1:(length(sales) + 4), ])  |>
  fmt_date(column = 2) |>
  fmt_currency(columns = 3, decimals = 0, currency = "USD") |>
  fmt_currency(columns = 4, decimals = 0, currency = "USD") |>
  fmt_currency(columns = 5, decimals = 1, currency = "USD") |>
  cols_label(sales_reported = "sales, thousands") |>
  cols_label(exponential_model = "exp, thousands") |>
  cols_label(yearly_rate_predicted_M = "annual, M") |>
  tab_header(
    title = md("Exponential Model"))

df_exp[1:(length(sales) + 4), 1:4] %>% 
  pivot_longer(3:4) %>% 
  ggplot(., aes(wk, value, colour = name)) + geom_point() + geom_line() +
  ylab("$, thousands") +
    ggtitle("Tonmya sales with exponential model")

df_exp[, 1:4] %>% 
  pivot_longer(3:4) %>% 
  ggplot(., aes(wk, value, colour = name)) + geom_point() + geom_line() +
  ylab("$, thousands") +
    ggtitle("Tonmya sales with exponential model", subtitle = "long extrapolation - for amusement?")

wk_cohort <- rep(as.character(1:4), 13)
df_exp$wk_cohort <- wk_cohort[1:length(df_exp$wk)]
```

I think we can see some trending on a monthly basis. Reporting doesn't match a monthly basis, but we can see if four-week intervals are interesting.

```{r sales cohorts}
#| echo: false
#| warning: false

df_exp[1:(length(sales)), c(1, 3, 5, 6)] %>% 
  ggplot(., aes(wk, sales_reported)) + 
  geom_point() + 
  geom_line() +
  ylab("$, thousands") +
    ggtitle("Tonmya sales with exponential model",
            "broken into week cohorts 1 through 4") +
  facet_grid(vars(wk_cohort))
#clean up
rm(df_exp)

```
## Sales Projection - at weekly rate - *linear model*

```{r sales linear model}
#| echo: false
#| warning: false

mod_sales_lin <- lm(sales ~ wk, data = df_sales)

summary(mod_sales_lin)
# plot(mod_sales_lin)

df_lin <- df_exp <- data.frame("wk" = 1:52, weekly_seq, 
           "sales_reported" = c(df_sales$sales, rep(NA, 52 - length(df_sales$sales))),
            "linear_model" = summary(mod_sales_lin)$coefficients[1] + summary(mod_sales_lin)$coefficients[2] * 1:52,
            "yearly_rate_predicted_M" = 52 / 1e3 * summary(mod_sales_lin)$coefficients[2] * 1:52)

gt(df_lin[1:(length(sales) + 4), ]) |>
  fmt_date(column = 2) |>
  fmt_currency(columns = 3, decimals = 0, currency = "USD") |>
  fmt_currency(columns = 4, decimals = 0, currency = "USD") |>
  fmt_currency(columns = 5, decimals = 1, currency = "USD") |>
  cols_label(sales_reported = "sales, thousands") |>
  cols_label(linear_model = "linear, thousands") |>
  cols_label(yearly_rate_predicted_M = "annual, M") |>
  tab_header(title = md("Linear Model"))

df_lin[1:(length(sales) + 4), 1:4] %>% 
  pivot_longer(3:4) %>% 
  ggplot(., aes(wk, value, colour = name)) + geom_point() + geom_line() +
  ylab("$, thousands") +
    ggtitle("Tonmya sales with linear model")

df_lin[, 1:4] %>% 
  pivot_longer(3:4) %>% 
  ggplot(., aes(wk, value, colour = name)) + geom_point() + geom_line() +
  ylab("$, thousands") +
    ggtitle("Tonmya sales with linear model", subtitle = "long extrapolation")

#clean up
rm(df_lin)

```

## Scripts Projection - at weekly growth rate - *exponential model*


```{r scripts exponential model}
#| echo: false
#| warning: false

df_scripts <- data.frame(scripts, "wk" = 1:length(scripts))
mod_scripts_exp <- nls(scripts ~ a * (1 + r)^wk,
           data = df_scripts,
           start = list(a = 17, r = 0))
summary(mod_scripts_exp)

df_exp <- data.frame("wk" = 1:52, weekly_seq, 
           "scripts_reported" = c(df_scripts$scripts, rep(NA, 52 - length(df_scripts$scripts))),
           "exponential_model" = round(summary(mod_scripts_exp)$coefficients[1] *
              (1 + summary(mod_scripts_exp)$coefficients[2])^(1:52)),
            "yearly_rate_predicted_M" = 52 / 1e3 * summary(mod_scripts_exp)$coefficients[1] *
              (1 + summary(mod_scripts_exp)$coefficients[2])^(1:52))

gt(df_exp[1:(length(scripts) + 4), ]) |>
  fmt_date(column = 2) |>
  fmt_number(columns = 5, decimals = 1) |>
  cols_label(scripts_reported = "total reported, thousands") |>
  cols_label(exponential_model = "exp predicted, thousands") |>
  cols_label(yearly_rate_predicted_M = "annual, M") |>
  tab_header(
    title = md("Scripts, Exponential Model"))


df_exp[1:(length(scripts) + 4), 1:4] %>% 
  pivot_longer(3:4) %>% 
  ggplot(., aes(wk, value, colour = name)) + geom_point() + geom_line() +
  ylab("total scripts") +
    ggtitle("Tonmya scripts with exponential model")
```

TRx Total Prescriptions by Weekly Cohorts

```{r scripts cohorts}
#| echo: false
#| warning: false

df_exp$wk_cohort <- wk_cohort[1:length(df_exp$wk)]
df_exp[1:(length(sales)), c(1, 3, 6)] %>% 
  ggplot(., aes(wk, scripts_reported)) + 
  geom_point() + 
  geom_line() +
  ylab("$, thousands") +
    ggtitle("Tonmya sales with exponential model",
            "broken into week cohorts 1 through 4") +
  facet_grid(vars(wk_cohort))
#clean up
rm(df_exp)
```

## scripts Projection - at weekly rate - *linear model*

```{r scripts linear model}
#| echo: false
#| warning: false

mod_scripts_lin <- lm(scripts ~ wk, data = df_scripts)

summary(mod_scripts_lin)
# plot(mod_scripts_lin)

df_lin <- data.frame("wk" = 1:52, weekly_seq, 
           "scripts_reported" = c(df_scripts$scripts, rep(NA, 52 - length(df_scripts$scripts))),
           "linear_model" = summary(mod_scripts_lin)$coefficients[1] + summary(mod_scripts_lin)$coefficients[2] * 1:52,
            "yearly_rate_predicted_M" = 52 / 1e3 * summary(mod_scripts_lin)$coefficients[2] * 1:52)

gt(df_lin[1:(length(scripts) + 4), ]) |>
  fmt_date(column = 2) |>
  fmt_number(columns = 4, decimals = 1) |>
  fmt_number(columns = 5, decimals = 1) |>
  cols_label(scripts_reported = "total reported, thousands") |>
  cols_label(linear_model = "linear predicted, thousands") |>
  cols_label(yearly_rate_predicted_M = "annual, M") |>
  tab_header(title = md("Scripts, Linear Model"))

df_lin[1:(length(scripts) + 4), 1:4] %>% 
  pivot_longer(3:4) %>% 
  ggplot(., aes(wk, value, colour = name)) + geom_point() + geom_line() +
  ylab("total scripts") +
    ggtitle("Tonmya scripts with linear model")

rm(df_lin)
```

## total / new scripts at weekly rate - *linear model*

```{r scripts ratio linear model}
#| echo: false
#| warning: false

trx_nrx_df <- data.frame("wk" = 1:length(scripts), "date" = weekly_seq[1:length(scripts)], "TRx" = scripts, "NRx" = new_scripts)
trx_nrx_df$total_to_new = trx_nrx_df$TRx / trx_nrx_df$NRx

mod_scripts_rat_lin <- lm(total_to_new ~ wk, data = trx_nrx_df)

summary(mod_scripts_rat_lin)
# plot(mod_scripts_lin)
trx_nrx_df$lin <- predict(mod_scripts_rat_lin, wk = data.frame(1:length(scripts)))

trx_nrx_df %>% 
  pivot_longer(5:6) %>% 
  ggplot(., aes(wk, value, colour = name)) + geom_point() + geom_line() +
  ylab("TRx/NRx") +
    ggtitle("Tonmya total / new prescriptions linear model")
```
